---
title: "Replication of Lee & Zhang (2013), OA 2" 
author: "Christian Mueller"
date: "3rd September 2015"
output: 
  html_document: 
    keep_md: yes
---

Validate the computation of Myers' index by replicating
an analysis, which appears in Online Appendix 2 of Lee and Zhang's
2013 research paper.

Melissa M. Lee and Nan Zhan (2013). The Art of Counting the Governed:
Census Accuracy, Civil War, and State Presence. CDDRL Working Papers, 
Vol. 146. Available [here](http://cddrl.fsi.stanford.edu/publications/the_art_of_counting_the_governed_census_accuracy_civil_war_and_state_presence).

## Prepare

```{r}
suppressPackageStartupMessages(source("../myers-score.R"))
```

```{r}
lz <- read.csv("../tests/data/lz-oa2.1.csv")
```

## Validate

Test if raw computation matches the result that Lee and Zhang compute.

```{r}
myers(lz$age, lz$freq, bin_start = 10, bin_size = 60)
```

Lee and Zhang (2013: Table OA2.3 on p. 40) report the correct result as 17.911%.

## Validate internals

Also validate each step in the computation using the data of Lee & Zhang.

```{r}
target <- lz$age
freq <- lz$freq
bin_start <- 10
bin_size <- 60

d <- data.frame(target = target, freq = freq)
d_filter <- seq(bin_start, bin_start + bin_size + 10 - 2)

d <- d %>%
  filter(target %in% d_filter) %>%
  arrange(target) %>%
  mutate(lastdigit = target %% 10, wgt = c(1:9, rep(10, bin_size - (10 - 1)), 9:1))
```

I just copy-pasted the definition of the function but I removed the `stop()` and the `na.omit()` calls because the are not needed here.

```{r}
group_by(d, lastdigit) %>%
  summarise(freq = sum(freq * wgt))
```

This computes the total terminal digits in the "blended" poulation
It matches column (k) in table OA 2.2 and column (5) in table OA 2.3.

```{r}
group_by(d, lastdigit) %>%
  summarise(freq = sum(freq * wgt)) %>%
  mutate(
    freq = freq / sum(freq),
    score = (freq * 100) - 10,
    score = abs(score)
  )
```

This computes the percent of each terminal digit in the "blended"
population as well as the deviation from the expected 10%.
It matches columns (6) and (7) of table OA 2.3.
